---
title: "YKKY0081天津胸科医院-个性化"
author: "GuoHao"
date: "2024/3/6"
output: html_document
---

## R Package loading

```{r load.package, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
suppressPackageStartupMessages(library(EPARS))
suppressPackageStartupMessages(library(ComplexHeatmap))
suppressPackageStartupMessages(library(openxlsx))
suppressPackageStartupMessages(library(RColorBrewer))
suppressPackageStartupMessages(library(corrplot))
suppressPackageStartupMessages(library(data.table))
suppressPackageStartupMessages(library(ggpubr))
suppressPackageStartupMessages(library(limma))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(pryr))
suppressPackageStartupMessages(library(tidyHeatmap))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(kableExtra))
suppressPackageStartupMessages(library(DT))
suppressPackageStartupMessages(library(clusterProfiler))
suppressPackageStartupMessages(library(org.Hs.eg.db))
suppressPackageStartupMessages(library(Hmisc))
suppressPackageStartupMessages(library(base64enc))
suppressPackageStartupMessages(library(R.utils))
suppressPackageStartupMessages(library(plotly))
suppressPackageStartupMessages(library(stringr))
suppressPackageStartupMessages(library(RColorBrewer))
suppressPackageStartupMessages(library(corrplot))
suppressPackageStartupMessages(library(data.table))
suppressPackageStartupMessages(library(ggpubr))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggthemes))
suppressPackageStartupMessages(library(ggsci))
suppressPackageStartupMessages(library(survival))
suppressPackageStartupMessages(library(survminer))
suppressPackageStartupMessages(library(CellChat))
suppressPackageStartupMessages(library(patchwork))
suppressPackageStartupMessages(library(SpatialDecon))
suppressPackageStartupMessages(library(RColorBrewer))
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
setwd("~/work/DSP/YKKY0081天津胸科医院/DSP_WTA")
# default parameter
Work_canche_dir <- "./personal_analysis/rmd_Results/"
Work_plot_dir<- "./personal_analysis/rmd_Results/"
```

## Creat Function

```{r set.the.function}
gh.subset <- function(object, patter) {
  object@anno <- object@anno %>%
    dplyr::filter({{ patter }})
  object@exp <- object@exp %>%
    as.data.frame() %>%
    dplyr::select(rownames(object@anno)) %>%
    as.matrix()
  return(object)
}

heatmap.function <- function(object, PlotsPath, i) {
  exp_group_object_single <- object
  exp_group_object_single@heatmap_groups <- c("CancerType", "Sample")
  exp_group_object_single <- epars_plot_setting(exp_group_object_single)
  exp_group_object_single <- epars_heatmap_anno_data(exp_group_object_single)
  rownames(exp_group_object_single@plot_setting$group_df) <- rownames(exp_group_object_single@anno)
  # group_df_order = order(exp_group_object_single@plot_setting$group_df[,1])
  # ordered_sample <- rownames(exp_group_object_single@plot_setting$group_df)[group_df_order]
  FigWidth <- 0.1 * nrow(exp_group_object@anno) + 3
  # 配置单分组的显示样式
  heat_mat <- htmp_scale(exp_group_object_single@exp)
  p <- Heatmap(heat_mat,
    show_column_names = FALSE,
    show_row_names = FALSE,
    heatmap_legend_param = list(title = "", legend_height = unit(4, "cm")),
    # column_order = ordered_sample,
    column_names_rot = -90,
    row_names_gp = gpar(fontsize = 4),
    column_names_gp = gpar(fontsize = 5),
    clustering_method_rows = "ward.D2",
    name = "Scale(exp)",
    cluster_column_slices = F,
    cluster_rows = T,
    cluster_columns = T,
    # column_split = exp_group_object_single@plot_setting$group_df,
    column_title = paste0(ncol(heat_mat), " ", i, "-AOIs"),
    column_title_side = "bottom",
    column_title_gp = gpar(fontsize = 15),
    heatmap_width = unit(5, "npc"),
    heatmap_height = unit(7, "npc"),
    clustering_method_columns = "ward.D2",
    top_annotation = exp_group_object_single@plot_setting$heatmap_anno,
    col = exp_group_object@config$heatmap_color
  )
  suppressMessages(save_pdf(p, PlotsPath, width = FigWidth))
}

ghboxplt <- function(data, patter) {
  data %>%
    dplyr::filter(Marker_names == {{ patter }}) %>%
    ggboxplot(
      x = "CancerType",
      y = "Marker_value",
      fill = "Morphology",
      ylab = paste0(patter, " Value")
    ) +
    ggsci::scale_fill_aaas()
}

List2ggupset <- function(list) {
  gene_pathway_membership <- fromList(list) %>%
    apply(2, as.logical)
  rownames(gene_pathway_membership) <- unique(unlist(list))

  tidy_pathway_member <- gene_pathway_membership %>%
    t() %>%
    as_tibble(rownames = "Pathway") %>%
    gather(Gene, Member, -Pathway) %>%
    filter(Member) %>%
    select(-Member) %>%
    group_by(Gene) %>%
    summarise(Pathways = list(Pathway))
  return(tidy_pathway_member)
}

upset.plot <- function(data, ...) {
  p <- data %>%
    ggplot(aes(x = Pathways)) +
    geom_bar() +
    scale_x_upset(
      order_by = "degree",
      reverse = T,
    ) +
    geom_text(stat = "count", aes(label = after_stat(count)), vjust = -1) +
    # scale_y_continuous(name = "Intersection size",expand = c(0, 0), limits = c(0, 6))+
    # theme(text = element_text(family = "ARIAL", colour = "black", size = 9))+
    xlab("") +
    theme_combmatrix(
      # combmatrix.label.text = element_text(family = "ARIAL", colour = "black", size = 9),
    ) +
    theme(panel.border = element_rect(colour = "black", fill = NA, size = 2)) +
    theme(
      panel.background = element_blank(),
      axis.text.y = element_text(size = 9)
    ) +
    ggtitle("Differential genes (control:case)")
  return(p)
}

enrichKEGG.function <- function(gene, qvalueCutoff = 1, pvalueCutoff = 1) {
  gene <- tranSymbol(gene)
  enr <- enrichKEGG(gene$ENTREZID, pvalueCutoff = pvalueCutoff, qvalueCutoff = qvalueCutoff)
  return(enr)
}

trans2num <- function(Ratio) {
  fractions <- sapply(strsplit(Ratio, "/"), as.integer)
  numeric_values <- fractions[1, ] / fractions[2, ]
  numeric_values <- round(numeric_values, digits = 2)
  return(numeric_values)
}

dotplot.function <- function(data) {
  p <- data %>%
    ggplot(aes(x = Group, y = Description)) +
    geom_point(aes(size = GeneRatio, color = p.adjust)) +
    theme_bw() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      # axis.line = element_line(colour = "black", linewidth = 1),
      panel.border = element_rect(color = "black", linewidth = 1, fill = NA),
      panel.grid.major = element_line(colour = "lightgrey")
    )
  return(p)
}

nordata <- function(data) {
  return(data / sum(data))
}
output_intersection <- function(intersecion_set, file){
  intersecion_set$Pathways <- as.character(intersecion_set$Pathways)
  intersecion_set <- split(intersecion_set, intersecion_set$Pathways) %>% 
    lapply(\(.x) dplyr::select(.x, Gene))
  names(intersecion_set) <-
    names(intersecion_set) %>% 
    gsub("c\\(", "", .) %>% 
    gsub(":", "vs", .) %>% 
    str_extract_all("\\((.*?)\\)") %>% 
    lapply(\(.x) paste0(.x, collapse = "_")) %>% 
    gsub("\\(|\\)", "", .) 
  intersecion_set <- lapply(names(intersecion_set), \(.x){
    intersecion_set[[.x]] <- intersecion_set[[.x]] %>% 
      dplyr::mutate(Group = .x)
  }) %>% 
    do.call(rbind, .)
  write.xlsx(intersecion_set, file = file)
  return(intersecion_set)
  
}

enrichtment_tidydata <- function(DEG_lis, CachePath){
  enrichment.res_list <- DEG_list %>% 
  purrr::map_depth(3, function(.x){
    rownames(.x) %>% 
      enrichKEGG.function(qvalueCutoff = 1, pvalueCutoff = 1)
  }) %>% # del the null list
  purrr::map_depth(2, function(.x){
    .x[!sapply(.x, \(.x) is.null(.x))]
  }) %>% 
  purrr::map_depth(1, function(.x){
    .x[sapply(.x, \(.x) length(.x) != 0)]
  })
saveRDS(enrichment.res_list, paste0(CachePath, "enrichment.res.RDs"))
enrichment.res_list <- 
  list(UP = enrichment.res_list %>% 
         lapply(\(.x){
           purrr::map(.x, ~ .x[["UP"]])
         }) %>% # del the null list
         purrr::map_depth(1, function(.x){
           .x[sapply(.x, \(.x) length(.x) != 0)]
         }),
       DOWN = enrichment.res_list %>% 
         lapply(\(.x){
           purrr::map(.x, ~ .x[["DOWN"]])
         }) %>% # del the null list
         purrr::map_depth(1, function(.x){
           .x[sapply(.x, \(.x) length(.x) != 0)]
         })
  ) %>% 
  purrr::map_depth(3, function(.x) .x@result)
return(enrichment.res_list)
}
### DSP_FindAllMarker
source("~/script/Rscript/DSP_FindAllMarker.r")
### Part V
mean_Patient <- function(meandata, object = exp_group_object){
  exp <- aggregate(meandata[, 12:ncol(meandata)], by=list(meandata$TMA.position),mean) #计算同一病人不同AOI的均值
  rownames(exp) <- exp[,1]
  exp <- exp[,-1]
  data <- object
  data@exp <- t(exp)
  data@anno <- meandata[, 1:11] %>% distinct(TMA.position, .keep_all = T)
  rownames(data@anno) <- data@anno$TMA.position
  data@anno$OS <- strsplit(data@anno$PFS_OS, " OS") %>% Reduce(rbind,.) %>% 
    as.data.frame() %>% .[,2]
  data@anno$DFS <- strsplit(data@anno$PFS_OS, " OS") %>% Reduce(rbind,.) %>% 
    as.data.frame() %>% .[,1]
  data@anno["08-1012-2",c("OS","DFS")] <- c("175","172.8")
  data@anno$OS <- data@anno$OS %>% gsub(pattern = " ",replacement = "")
  data@anno$DFS <- data@anno$DFS %>% gsub(pattern = "DFS ",replacement = "") %>% 
    gsub(pattern = "DFS",replacement = "")
  data@anno["18-2116-4","DFS"] <- "54.8"
  data@anno$Status <- data@anno$Status %>% gsub(pattern = "未复发",replacement = "0") %>% 
    gsub(pattern = "复发",replacement = "1") %>% as.numeric()
  data@anno$OS <- as.numeric(data@anno$OS)
  data@anno$DFS <- as.numeric(data@anno$DFS)
  return(data)
}
log_P <- function(data, intersect_gene = all_intersect_gene){
  # browser()
  exp <- data@exp %>% 
    t() %>% 
    as.data.frame() %>% 
    dplyr::select(unlist(intersect_gene) %>% as.character() %>% unique())
  anno <- data@anno %>% 
    dplyr::select(DFS, Status)
  cal_data <- merge(anno, exp, by = 0) %>% 
    column_to_rownames("Row.names")
  phe <- cal_data[, 1:2]
  mySurv <- with(cal_data, Surv(DFS, Status))
  exprSet <- cal_data[, -c(1:2)] %>% t
  log_rank_p <- apply(exprSet, 1 , function(gene){
    # gene=exprSet[1,]
    phe$group <- ifelse(gene > median(gene), 'high', 'low')  
    data.survdiff <- survdiff(mySurv ~ group, data = phe)
    p.val = 1 - pchisq(data.survdiff$chisq, length(data.survdiff$n) - 1)
    return(p.val)
  })
  log_rank_p <- sort(log_rank_p)
  # table(log_rank_p < 0.05)
  sig_gene <- log_rank_p[log_rank_p < 0.05] %>% names()
  return(sig_gene)
}
SurPlot <- function(data, sig_gene, listnames = "TC", CachePath) {
  # browser()
  exprSet <- data[[listnames]]@exp %>%
    as.data.frame() %>%
    dplyr::filter(rownames(.)  %in% sig_gene[[listnames]]) %>%
    as.matrix()
  rownames(exprSet) <-
    gsub(rownames(exprSet),
         pattern = "-",
         replacement = "_")
  # phe <- data[[listnames]]@anno %>%
  #   dplyr::select(DFS, Status)
  for (g in gsub("-", "_", sig_gene[[listnames]])) {
    phe <- data[[listnames]]@anno %>%
      dplyr::select(DFS, Status)
    phe <- phe %>%
      dplyr::mutate(!!sym(g) := ifelse(exprSet[g,] > median(exprSet[g,]), 'high', 'low'))
    print(g)
    # a <- as.formula(paste0("Surv(DFS, Status) ~ ", g))
    # sfit1 <- survfit(!!a, data = phe)
    formula_str <- paste("Surv(DFS, Status)", g, sep = " ~ ")
    formula_str <- as.formula(formula_str)
    sfit1 <- survfit(formula_str, data = phe)
    sfit1[["call"]][["formula"]] <- formula_str
    print(sfit1[["call"]][["formula"]])
    p <-
      ggsurvplot(sfit1,
                 pval = TRUE,
                 data = phe,
                 risk.table = TRUE)
    pdf(file = paste0(CachePath, g, "_survplot.pdf"),
        onefile = F)
    print(p)
    dev.off()
  }
}
DSP_subset <- function(object, patter = NULL){
  object@anno <- object@anno %>% 
    dplyr::filter(!!rlang::parse_expr(patter))
  object@raw_exp <- object@raw_exp %>% 
    as.data.frame() %>% 
    dplyr::select(all_of(rownames(object@anno))) %>% as.matrix()
  object@exp <- object@exp %>% 
    as.data.frame() %>% 
    dplyr::select(all_of(rownames(object@anno))) %>% as.matrix()
  return(object)
}
SpatialDecon_summary <- function(object){
  per.observation.mean.neg <- object@exp["NegProbe-WTX", ]
  bg <-  sweep(object@exp * 0, 2, per.observation.mean.neg, "+")
  bg2 <- derive_GeoMx_background(norm = object@exp,
                                 probepool = rep(1, nrow(object@exp)),
                                 negnames = "NegProbe-WTX")
  res <- spatialdecon(norm = object@exp, bg = bg, X = safeTME,
                      cell_counts = object@anno$AOINucleiCount,
                      # cellmerges = safeTME.matches, # 是否合并成大类
                      # is_pure_tumor = object@anno$Tumor == "TRUE",
                      align_genes = TRUE)
  return(res)
}
```

## 样本信息

### 分组

Group1: 恶性程度从低到高：贴壁型（LPA）、腺泡型（APA）、乳头型（PPA）、复杂腺体型（CAA）、实性型（SPA）、微乳头型（MPA） Group2: 预后分组：预后好 vs 预后差 \#\#\# 分析需求 以下所有分析去掉1个AOI（16-2060-1-MPA-BY-Full）

```{r del.the.sample}
## load the raw data, the data neet to del the sample
load("./results/rendering_cache_DSP_CTA/data-readin_3dab01df010760cdad434e6c2cae1f1c.RData")
exp_group_object@anno <- exp_group_object@anno %>%
  dplyr::filter(rownames(.) != "23R3921.3843.7918SLZA_007_Full")
exp_group_object@exp <- exp_group_object@exp %>%
  as.data.frame() %>%
  dplyr::select(-`23R3921.3843.7918SLZA_007_Full`) %>%
  as.matrix()
exp_group_object@anno <- exp_group_object@anno %>%
  mutate(Sample = TMA.position)
saveRDS(exp_group_object, file = "./personal_analysis/rmd_Results/exp_group_object.RDs")
rm(list = ls())
```

## Part I of the article

### 1.1 dimensionality reduction

#### Integral dimension reduction

```{r integral.Integral.dimension, warning=FALSE}
exp_group_object <- readRDS("./personal_analysis/rmd_Results/exp_group_object.RDs")
CachePath <- paste0(Work_canche_dir, "Part I/1.1 dimensionality reduction/integral level/")
PlotsPath <- paste0(Work_canche_dir, "Part I/1.1 dimensionality reduction/integral level/")
create_dir(path.list = c(CachePath, PlotsPath))
exp_group_object@heatmap_groups <- c("Morphology", "TNM", "CancerType", "Prognosis", "Sample")
exp_group_object <- PCA_plot(exp_group_object,
  CachePath = CachePath,
  group = exp_group_object@heatmap_groups,
  color_palette = exp_group_object@config$pan_color,
  template = "", space = " ",
  prefix = "PCA"
)
exp_group_object <- tSNE_plot(exp_group_object,
  CachePath = CachePath,
  group = exp_group_object@heatmap_groups,
  color_palette = exp_group_object@config$pan_color,
  template = "", space = "",
  prefix = "tSNE"
)
exp_group_object <- UMAP_plot(exp_group_object,
  CachePath = CachePath,
  group = exp_group_object@heatmap_groups,
  color_palette = exp_group_object@config$pan_color,
  template = "", space = "",
  prefix = "UMAP"
)
```

#### Group1 降维

```{r Group1}
CachePath <- paste0(Work_canche_dir, "Part I/1.1 dimensionality reduction/Group1/placeholder/")
PlotsPath <- paste0(Work_canche_dir, "Part I/1.1 dimensionality reduction/Group1/placeholder/")
for (i in unique(exp_group_object@anno$Morphology)) {
  CachePath <- sub("[/][^/]+$", paste0("/", i, "/"), gsub("[/]$", "", CachePath))
  PlotsPath <- sub("[/][^/]+$", paste0("/", i, "/"), gsub("[/]$", "", PlotsPath))
  # print(CachePath)
  create_dir(path.list = c(CachePath, PlotsPath))
  object <- exp_group_object %>%
    gh.subset(Morphology == i)
  object@heatmap_groups <- c("CancerType")
  object <- PCA_plot(object,
    CachePath = CachePath,
    group = object@heatmap_groups,
    color_palette = object@config$pan_color,
    template = "", space = " ",
    prefix = paste0("PCA_", i)
  )
  object <- tSNE_plot(object,
    CachePath = CachePath,
    group = object@heatmap_groups,
    color_palette = object@config$pan_color,
    template = "", space = "",
    prefix = paste0("tSNE_", i)
  )
  object <- UMAP_plot(object,
    CachePath = CachePath,
    group = object@heatmap_groups, n_neighbors = 6,
    color_palette = object@config$pan_color,
    template = "", space = "",
    prefix = paste0("UMAP_", i)
  )
  rm(object)
}
### point 5
CachePath <- paste0(Work_canche_dir, "Part I/1.1 dimensionality reduction/Group1/CancerType/placeholder/")
PlotsPath <- paste0(Work_canche_dir, "Part I/1.1 dimensionality reduction/Group1/CancerType/placeholder/")
# create_dir(path.list = c(CachePath,PlotsPath))
for (i in unique(exp_group_object@anno$CancerType)) {
  CachePath <- sub("[/][^/]+$", paste0("/", i, "/"), gsub("[/]$", "", CachePath))
  PlotsPath <- sub("[/][^/]+$", paste0("/", i, "/"), gsub("[/]$", "", PlotsPath))
  # print(CachePath)
  create_dir(path.list = c(CachePath, PlotsPath))
  object <- exp_group_object %>%
    gh.subset(CancerType == i)
  object@heatmap_groups <- c("Prognosis")
  object <- PCA_plot(object,
    CachePath = CachePath,
    group = object@heatmap_groups,
    color_palette = object@config$pan_color,
    template = "", space = " ",
    prefix = paste0("PCA_", i)
  )
  object <- tSNE_plot(object,
    CachePath = CachePath,
    group = object@heatmap_groups,
    color_palette = object@config$pan_color,
    template = "", space = "",
    prefix = paste0("tSNE_", i)
  )
  object <- UMAP_plot(object,
    CachePath = CachePath,
    group = object@heatmap_groups, n_neighbors = 6,
    color_palette = object@config$pan_color,
    template = "", space = "",
    prefix = paste0("UMAP_", i)
  )
  rm(object)
}
```

#### Group2 降维

```{r Group2 }
CachePath <- paste0(Work_canche_dir, "Part I/1.1 dimensionality reduction/Group2/")
PlotsPath <- paste0(Work_canche_dir, "Part I/1.1 dimensionality reduction/Group2/")
create_dir(path.list = c(CachePath, PlotsPath))
for (i in unique(exp_group_object@anno$Morphology)) {
  object <- exp_group_object %>% 
    gh.subset(Morphology == i)
  object@heatmap_groups <- c("Prognosis")  
  object <- PCA_plot(object, CachePath = CachePath,
                     group = object@heatmap_groups, 
                     color_palette = object@config$pan_color,
                     template = "", space = " ",
                     prefix = paste0("PCA_", i))
  object <- tSNE_plot(object, CachePath = CachePath,
                      group = object@heatmap_groups,
                      color_palette = object@config$pan_color,
                      template = "", space = "",
                      prefix = paste0("tSNE_", i))
  object <- UMAP_plot(object, CachePath = CachePath,
                      group = object@heatmap_groups, n_neighbors = 6,
                      color_palette = object@config$pan_color,
                      template = "", space = "",
                      prefix = paste0("UMAP_", i))
  rm(object)
} 
```

### 1.2 Overview HeatMap

```{r Overview_Heatmap}
CachePath <- paste0(Work_canche_dir, "Part I/1.2 Overview Heatmap/")
PlotsPath <- paste0(Work_canche_dir, "Part I/1.2 Overview Heatmap/")
create_dir(path.list = c(CachePath, PlotsPath))
exp_group_object_single <- exp_group_object
# exp_group_object_single@exp %>% rownames() %>% grep(pattern = "WTX", value = T)
exp_group_object_single@exp <- exp_group_object_single@exp %>% 
  as.data.frame() %>% 
  dplyr::filter(!grepl("WTX", rownames(.))) %>% 
    as.matrix()

for (i in unique(exp_group_object@anno$Morphology)) {
  object <- exp_group_object_single %>% 
    gh.subset(Morphology == i)
  # debug(heatmap.function)
  heatmap.function(object, paste0(PlotsPath, "_", i, ".pdf"), i)
}
rm(exp_group_object_single)
```

### 1.3 Mark gene expression

```{r Mark gene expression}
CachePath <- paste0(Work_canche_dir, "Part I/1.3 Mark gene expression/")
PlotsPath <- paste0(Work_canche_dir, "Part I/1.3 Mark gene expression/")
create_dir(path.list = c(CachePath, PlotsPath))

panck.marker <- c("KRT1", "KRT5", "KRT8", "KRT14", "KRT19")
plotdata <- exp_group_object@exp %>% 
  as.data.frame() %>% 
  dplyr::filter(rownames(.) %in% c(panck.marker, "CD68", "CD163")) %>% 
  t %>% 
  as.data.frame() %>% 
  merge(exp_group_object@anno %>% 
          dplyr::select(Morphology, Sample, CancerType), by = 0, sort = F) %>% 
  rename(SampleID = Row.names) %>% 
  mutate(PanCK = (KRT8+KRT19)/2) %>% # 只能找到这两个基因, 大抵是被过滤掉了
  dplyr::select(-KRT19, -KRT8) %>% 
  dplyr::select(1, 4, 5, 6, everything()) %>% 
  pivot_longer(cols = -colnames(.)[1:4], names_to = "Marker_names", values_to = "Marker_value")

lapply(unique(plotdata$Marker_names), function(.x){
  p <- ghboxplt(plotdata, .x)
  ggsave(p, filename = paste0(PlotsPath, .x, "boxplot.pdf"))
})
```

## Part II of the article

### 2.1 Search for lung tumor markers in epithelial cells

首先进行差异基因的计算, 将所有的亚型跟最恶性程度最低的亚型LPA相比, 寻找差异基因, 并取交集.

这里一并计算了M1/M2中的差异基因备用. 本部分筛选仍使用上皮细胞相关的数据.

没有找到再多个差异分组中都显著的基因, 也没有找到随着恶性程度上升表达量上升的基因, 随之下降的也没有.

```{r DEG}
CachePath <- paste0(Work_canche_dir, "Part II/2.1 differential gene/")
PlotsPath <- paste0(Work_canche_dir, "Part II/2.1 differential gene/")
create_dir(path.list = c(CachePath, PlotsPath))
exp_group_object@deg_groups <- c("CancerType(LPA:PPA),Morphology$M1; CancerType(LPA:CAA),Morphology$M1; CancerType(LPA:SPA),Morphology$M1; CancerType(LPA:MPA),Morphology$M1; CancerType(LPA:APA),Morphology$M2; CancerType(LPA:CAA),Morphology$M2; CancerType(LPA:SPA),Morphology$M2; CancerType(LPA:MPA),Morphology$M2; CancerType(LPA:APA),Morphology$TC; CancerType(LPA:PPA),Morphology$TC; CancerType(LPA:CAA),Morphology$TC; CancerType(LPA:SPA),Morphology$TC; CancerType(LPA:MPA),Morphology$TC")
source("~/script/Rscript/Parallel script/library_EPARS_parallel.r")
exp_group_object <- parallel_DEG(exp_group_object, CachePath = "./personal_analysis/rmd_Results/Part II/2.1 differential gene/DEG_res/", core.num = 6)

parallel_heatmap(exp_group_object, CachePath = "./personal_analysis/rmd_Results/Part II/2.1 differential gene/DEG_res/heatmap/", PlotsPath = "./personal_analysis/rmd_Results/Part II/2.1 differential gene/DEG_res/heatmap//", core.num = 6, top = 10)

parallel_volcano(exp_group_object, PlotsPath = "./personal_analysis/rmd_Results/Part II/2.1 differential gene/DEG_res/volcano_lplot/")


exp_list <- exp_group_object@deg %>% 
  lapply(\(.x){.x@result_df}) %>% 
  lapply(\(.x){
    .x %>% dplyr::filter(change != "NOT")  %>% 
      split(f = .$change)
  }) %>% 
  .[sapply(., \(.x) length(.x) != 0)] %>% 
  .[!grepl("^Pro", names(.))]

DEG_list <- list(M1_list =  exp_list[names(exp_list) %>% grep(patter = "M1", value = T)],
                 M2_list = exp_list[names(exp_list) %>% grep(patter = "M2", value = T)],
                 TC_list = exp_list[names(exp_list) %>% grep(patter = "TC", value = T)]) %>% 
  lapply(\(.x) {
    names(.x) <- names(.x) %>% str_extract(".*(?=\\,)")
    return(.x)
  })

## 单独取TC的差异基因计算交集
Reduce(intersect, DEG_list$TC_list) # 没有在各个组内都有的基因
```

#### 2.1.1 Upset

由于差异没有什么交集结果, 准备利用upset图来看看, 有哪些群之间有交集

```{r Upset}
library(UpSetR)
library(ComplexUpset)
library(ggupset)
names(DEG_list$TC_list)
lapply(names(DEG_list), \(.x){
  dir <- paste0(Work_plot_dir, "Part II/2.1 differential gene/2.1.1 UPset_plot/", .x, "/")
  dir.create(dir, recursive = T, showWarnings = F)
  data <- DEG_list[[.x]]
  intersecion_set <- List2ggupset(data)
  output_intersection(intersecion_set, file = paste0(dir, "instersection_gene.xlsx"))
  p <- upset.plot(intersecion_set)
  ggsave(plot = p, filename = paste0(dir, "upset_plot.pdf"), height = 9, width = 16)
})
```

通过upset图可以看到, ***CDC25B*****,** ***CH25H***, ***SMAD1*** 是在三个亚型中共有的差异基因, 因此绘制折线图呈现这三个基因的表达趋势,

发现这三个基因会在除SPA, MPA的亚型中是上升趋势.

```{r 3gene boxplot}
time.point <- c("LPA", "APA", "PPA", "CAA", "SPA", "MPA")
exp_group_object@anno$CancerType <- factor(exp_group_object@anno$CancerType, levels = time.point)
exp <- exp_group_object@exp %>% 
  as.data.frame() %>% 
  filter(rownames(.) %in% c("TMEM221", "XAGE1A")) %>% t
anno <- exp_group_object@anno %>% 
  dplyr::select(CancerType)
exp <- merge(anno, exp, by = 0) %>%
  dplyr::select(-1) %>% 
  dplyr::summarise(across(everything(), mean), .by = CancerType) %>% 
  pivot_longer(cols = -CancerType)  
exp$CancerType <- factor(exp$CancerType, levels = time.point)
p <- ggline(exp, x = "CancerType", y = "value", color = "name") 
ggsave(filename = paste0(Work_plot_dir, "Part II/3_gene.pdf"), plot = p)
```

由此得到, 是否SPA和MPA中基因总体表达是变低的?

```{r all_gene_boxplot}
exp <- exp_group_object@exp %>% 
  as.data.frame() %>% t 
exp <- merge(anno, exp, by = 0) %>%
  dplyr::select(-1) %>% 
  summarise(across(everything(), mean), .by = CancerType) %>% 
  pivot_longer(cols = -CancerType)
exp <- exp %>% 
  mutate(value2 = log2(value))
exp$CancerType <- factor(exp$CancerType, levels = time.point)
p <- ggboxplot(exp, x = "CancerType", y = "value2")
ggsave(filename = paste0(Work_plot_dir, "Part II/all_gene_boxplot.pdf"), plot = p)
```

发现确实两个亚型的整体基因表达是较低的.

因此建议去掉这两个亚型看是否有什么发现

### 2.2 Enrichment analysis

富集分析同样没有发现共有的通路, 但是还是按照要求分析一遍

```{r Enrichment analysis}
CachePath <- paste0(Work_canche_dir, "Part II/2.2 Enrichment analysis/")
PlotsPath <- paste0(Work_canche_dir, "Part II/2.2 Enrichment analysis/")
create_dir(path.list = c(CachePath, PlotsPath))
enrichment.res_list <- DEG_list %>% 
  purrr::map_depth(3, function(.x){
    rownames(.x) %>% 
      enrichKEGG.function(qvalueCutoff = 1, pvalueCutoff = 1)
  }) %>% # del the null list
  purrr::map_depth(2, function(.x){
    .x[!sapply(.x, \(.x) is.null(.x))]
  }) %>% 
  purrr::map_depth(1, function(.x){
    .x[sapply(.x, \(.x) length(.x) != 0)]
  })
saveRDS(enrichment.res_list, paste0(CachePath, "enrichment.res.RDs"))
enrichment.res_list <- readRDS("~/work/DSP/YKKY0081天津胸科医院/DSP_WTA/personal_analysis/rmd_Results/Part II/2.2 Enrichment analysis/enrichment.res.RDs")
enrichment.res_list <- 
  list(UP = enrichment.res_list %>% 
            lapply(\(.x){
            purrr::map(.x, ~ .x[["UP"]])
            }) %>% # del the null list
            purrr::map_depth(1, function(.x){
              .x[sapply(.x, \(.x) length(.x) != 0)]
            }),
          DOWN = enrichment.res_list %>% 
            lapply(\(.x){
              purrr::map(.x, ~ .x[["DOWN"]])
            }) %>% # del the null list
            purrr::map_depth(1, function(.x){
              .x[sapply(.x, \(.x) length(.x) != 0)]
            })
          ) %>% 
  purrr::map_depth(3, function(.x) .x@result)
plot.data <- enrichment.res_list %>% 
  purrr::map_depth(2, function(.x){
    do.call(rbind, .x) %>% 
      rownames_to_column("Group") %>% 
      mutate(Group = str_extract(Group, ".*(?=\\.)")) %>% 
      group_by(Group) %>% # 保存表格注释掉
      slice_min(p.adjust, n = 5) %>% # 这里取了每个组中矫正p值前5位 # 保存表格注释掉
      # 在画图之前, 先保存字符形式的Ratio
      mutate(GeneRatio = trans2num(GeneRatio), BgRatio = trans2num(BgRatio)) # 保存表格注释掉
  })
lapply(names(plot.data), \(.x){
  lapply(names(plot.data[[.x]]), \(..x){
    dir <- paste0(CachePath, "DEG/Prognosis_enrichment_res/", .x, "/", ..x, "/")
    dir.create(dir, recursive = T, showWarnings = F)
    write.xlsx(plot.data[[.x]][[..x]], file = paste0(dir, "enrichment.results.xlsx"))
  })
})
lapply(names(plot.data), function(.x){
  lapply(names(plotdata[[.x]]), \(..x){
    print(..x)
    dir <- paste0(CachePath, .x, "/", ..x, "/")
    dir.create(dir, recursive = T)
    p <- dotplot.function(plot.data[[.x]][[..x]])
    ggsave(filename = paste0(dir, "dotplot.pdf"), height = 20, width = 10, plot = p)
  })
})
```

### 2.3 Use SCORPIUS to simulate developmental trajectories

看起来也没有什么趋势

```{r SCORPIUS}
library(SCORPIUS)
group_names <- exp_group_object@anno %>% 
  dplyr::select(CancerType)
exp <- merge(group_names, exp_group_object@exp %>% t, by = 0) %>% 
  dplyr::select(-1,-2) %>% 
  as.matrix()
group_names <- group_names$CancerType %>% factor(levels = time.point)
space <- reduce_dimensionality(exp, "spearman")
draw_trajectory_plot(space, group_names, contour = TRUE)
traj <- infer_trajectory(space)
p <- draw_trajectory_plot(space, group_names, traj$path, contour = TRUE)
ggsave(filename = paste0(Work_plot_dir, "Part II/2.3 SCORPIUS/developmental trajectories.pdf"), plot = p)
```

## Part III Overall description of immune cells

### 3.1 SpatialDecon immune cells

#### 3.1.2 all sample immune cells analysis

This chunk to analysis immunoinfiltration use all sample.

```{r SpatialDecon}
CachePath <- paste0(Work_canche_dir, "Part III/")
PlotsPath <- paste0(Work_canche_dir, "Part III/")
create_dir(path.list = c(CachePath, PlotsPath))
per.observation.mean.neg <- exp_group_object@exp["NegProbe-WTX", ]
bg <-  sweep(exp_group_object@exp * 0, 2, per.observation.mean.neg, "+")
bg2 <- derive_GeoMx_background(norm = exp_group_object@exp,
                               probepool = rep(1, nrow(exp_group_object@exp)),
                               negnames = "NegProbe-WTX")
res <- spatialdecon(norm = exp_group_object@exp, bg = bg, X = safeTME,
                    cell_counts = exp_group_object@anno$AOINucleiCount,
                    # cellmerges = safeTME.matches, # 是否合并成大类
                    # is_pure_tumor = object@anno$Tumor == "TRUE",
                    align_genes = TRUE)
saveRDS(res, file = paste0(CachePath, "3.1 immune cells/spatialdecon.res.abundance.RDs"))
write.xlsx(res$beta %>% as.data.frame(),
           file = paste0(CachePath, "3.1 immune cells/spatialdecon.res.abundance.xlsx"), 
           rowNames = T, overwrite = T)

### del the CAA sample
exp_group_object_partIII <- exp_group_object
exp_group_object_partIII@anno <- exp_group_object_partIII@anno %>% 
  dplyr::filter(CancerType != "CAA")
exp_group_object_partIII@exp <- exp_group_object_partIII@exp %>% 
  as.data.frame() %>% 
  dplyr::select(rownames(exp_group_object_partIII@anno)) %>% 
  as.matrix()
exp_group_object_partIII@raw_exp <- exp_group_object_partIII@raw_exp %>% 
  as.data.frame() %>% 
  dplyr::select(rownames(exp_group_object_partIII@anno)) %>% 
  as.matrix()

per.observation.mean.neg <- exp_group_object_partIII@exp["NegProbe-WTX", ]
bg <-  sweep(exp_group_object_partIII@exp * 0, 2, per.observation.mean.neg, "+")
bg2 <- derive_GeoMx_background(norm = exp_group_object_partIII@exp,
                               probepool = rep(1, nrow(exp_group_object_partIII@exp)),
                               negnames = "NegProbe-WTX")
res <- spatialdecon(norm = exp_group_object_partIII@exp, bg = bg, X = safeTME,
                    cell_counts = exp_group_object_partIII@anno$AOINucleiCount,
                    # cellmerges = safeTME.matches, # 是否合并成大类
                    # is_pure_tumor = object@anno$Tumor == "TRUE",
                    align_genes = TRUE)
write.xlsx(res$beta %>% as.data.frame(),
           file = paste0(CachePath, "3.1 immune cells/del_CAA/spatialdecon.res.abundance.xlsx"), 
           rowNames = T, overwrite = T)
```

```{r plot}
exp_group_object@heatmap_groups <- c("CancerType", "SegmentLabel")
exp_group_object@anno$CancerType <- factor(exp_group_object@anno$CancerType, levels = time.point)
exp_group_object@anno$SegmentLabel <- factor(exp_group_object@anno$SegmentLabel)
exp_group_object <- epars_plot_setting(exp_group_object)
exp_group_object <- epars_heatmap_anno_data(exp_group_object)
ordersample <- rownames(dplyr::arrange(exp_group_object@anno, CancerType))
# abundance 
p <- Heatmap(matrix = res$beta %>% scale(), 
        top_annotation = exp_group_object@plot_setting$heatmap_anno, 
        cluster_columns = F, column_order = ordersample,
        heatmap_legend_param = list(title = "",legend_height = unit(4, "cm")),
        column_split = exp_group_object@plot_setting$group_df[, "SegmentLabel"])
save_pdf(p, 
         filename = paste0(CachePath, "3.1 immune cells//SpatialDecon.immcell.abundance.heatmap.pdf"), 
         height = 12, width = 12)
Cancertype.abundance <- res$beta %>% 
  as.data.frame() %>% 
  t() %>% 
  merge(exp_group_object@anno %>% dplyr::select(CancerType), ., by = 0) %>% 
  column_to_rownames("Row.names") %>% 
  group_by(CancerType) %>% 
  dplyr::summarise(dplyr::across(everything(), sum)) %>% 
  column_to_rownames("CancerType") %>% 
  apply(1, \(.x) .x/sum(.x)) %>% 
  as.data.frame() %>% 
  rownames_to_column("CellType") %>% 
  pivot_longer(cols = -CellType, names_to = "CancerType", values_to = "Cell.Proportion")
suppressPackageStartupMessages(library(RColorBrewer))
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))

Cancertype.abundance$CancerType <- factor(Cancertype.abundance$CancerType, levels = time.point)
  
p <- Cancertype.abundance %>% 
  ggplot(aes(x = CancerType, y = Cell.Proportion)) + 
  geom_col(aes(fill = CellType)) + 
  scale_fill_manual(values = col_vector) + 
  ylab("Cell Proportion")
ggsave(plot = p, filename = paste0(CachePath,"3.1 immune cells/SpatialDecon.immcell.proportion.pdf"), 
       height = 12, width = 12)
#### 
# 经过检查, monocytes.C 这个在某些亚型中丰度为0
#### 

#### immcell abundance boxplot ####
plotdata <- res$beta %>% 
  as.data.frame() %>% 
  t() %>% 
  merge(exp_group_object@anno %>% dplyr::select(CancerType), ., by = 0) %>% 
  dplyr::select(-monocytes.C)

plotdata$CancerType <- plotdata$CancerType %>% as.character()
source("~/script/Rscript/all.function.R")

adjusted_boxplot_nojitter(input_df = plotdata, 
                          id_colnames = colnames(plotdata)[1:2], 
                          x_aes = colnames(plotdata)[2], 
                          plot_dir = paste0(CachePath, "3.1 immune cells/boxplot/"),
                          plot_suffix = ".abundance.boxplot.pdf", 
                          comparisons_list = combn(unique(plotdata$CancerType), 2, simplify = F), 
                          x_order = time.point, 
                          plot_width = 25, 
                          ncol = 6,
                          # stat_method = "t.test", 
                          adjust_method = "BH")
```

```{r plot_del_CAA}
time.point <- c("LPA", "APA", "PPA", "SPA", "MPA")
exp_group_object_partIII@heatmap_groups <- c("CancerType", "SegmentLabel")
exp_group_object_partIII@anno$CancerType <- factor(exp_group_object_partIII@anno$CancerType, levels = time.point)
exp_group_object_partIII@anno$SegmentLabel <- factor(exp_group_object_partIII@anno$SegmentLabel)
exp_group_object_partIII <- epars_plot_setting(exp_group_object_partIII)
exp_group_object_partIII <- epars_heatmap_anno_data(exp_group_object_partIII)
ordersample <- rownames(dplyr::arrange(exp_group_object_partIII@anno, CancerType))
# abundance 
p <- Heatmap(matrix = res$beta %>% scale(), 
        top_annotation = exp_group_object_partIII@plot_setting$heatmap_anno, 
        cluster_columns = F, column_order = ordersample,
        heatmap_legend_param = list(title = "",legend_height = unit(4, "cm")),
        column_split = exp_group_object_partIII@plot_setting$group_df[, "SegmentLabel"])
save_pdf(p, 
         filename = paste0(CachePath, "3.1 immune cells/del_CAA/SpatialDecon.immcell.abundance.heatmap.pdf"), 
         height = 12, width = 12)
Cancertype.abundance <- res$beta %>% 
  as.data.frame() %>% 
  t() %>% 
  merge(exp_group_object_partIII@anno %>% dplyr::select(CancerType), ., by = 0) %>% 
  column_to_rownames("Row.names") %>% 
  group_by(CancerType) %>% 
  dplyr::summarise(dplyr::across(everything(), sum)) %>% 
  column_to_rownames("CancerType") %>% 
  apply(1, \(.x) .x/sum(.x)) %>% 
  as.data.frame() %>% 
  rownames_to_column("CellType") %>% 
  pivot_longer(cols = -CellType, names_to = "CancerType", values_to = "Cell.Proportion")

Cancertype.abundance$CancerType <- factor(Cancertype.abundance$CancerType, levels = time.point)
  
p <- Cancertype.abundance %>% 
  ggplot(aes(x = CancerType, y = Cell.Proportion)) + 
  geom_col(aes(fill = CellType)) + 
  scale_fill_manual(values = col_vector) + 
  ylab("Cell Proportion")
ggsave(plot = p, filename = paste0(CachePath,"3.1 immune cells/del_CAA/SpatialDecon.immcell.proportion.pdf"), 
       height = 12, width = 12)
#### 
# 经过检查, monocytes.C 这个在某些亚型中丰度为0
#### 

#### immcell abundance boxplot ####
plotdata <- res$beta %>% 
  as.data.frame() %>% 
  t() %>% 
  merge(exp_group_object_partIII@anno %>% dplyr::select(CancerType), ., by = 0) %>% 
  dplyr::select(-monocytes.C)

plotdata$CancerType <- plotdata$CancerType %>% as.character()
suppressWarnings(source("~/script/Rscript/all.function.R"))

adjusted_boxplot_nojitter(input_df = plotdata, 
                          id_colnames = colnames(plotdata)[1:2], 
                          x_aes = colnames(plotdata)[2], 
                          plot_dir = paste0(CachePath, "3.1 immune cells/del_CAA/boxplot/"),
                          plot_suffix = ".abundance.boxplot.pdf", 
                          comparisons_list = combn(unique(plotdata$CancerType), 2, simplify = F), 
                          x_order = time.point, 
                          plot_width = 25, 
                          ncol = 6,
                          # stat_method = "t.test", 
                          adjust_method = "BH")
rm(p, plotdata, res, bg2, bg, exp_group_object_partIII, Cancertype.abundance)
```

#### 3.1.3 Immunoinflitration of diiffernet prognostic samples

This chunk analysis immunoinfiltration of different prognositc samples

```{r prepare_the_data}
# CachePath <- paste0(Work_canche_dir, "Part III/3.1 immune cells/different_prognositc/")
# PlotsPath <- paste0(Work_canche_dir, "Part III/3.1 immune cells/different_prognositc/")
# create_dir(path.list = c(CachePath, PlotsPath))

# object_list <- list(Good = DSP_subset(exp_group_object, patter = "Prognosis == 'bad'"), 
#                     Bad = DSP_subset(exp_group_object, patter = "Prognosis == 'bad'"))
```

```{r calcultaion_immcellproportion}
# res <- lapply(object_list, \(.x) SpatialDecon_summary(.x))
# outputxlsx <- lapply(res, \(.x) as.data.frame(.x$beta))
# write.xlsx(outputxlsx, 
#            file = paste0(CachePath, "3.1 immune cells/different_prognositc/SpatialDecon_prognositc_res.xlsx"),
#            rowNames = T, overwrite = T)
### There is no need to calculate separately
res <- readRDS(file = paste0(CachePath, "3.1 immune cells/spatialdecon.res.abundance.RDs"))
```

```{r plot}
exp_group_object@heatmap_groups <- c("CancerType", "Prognosis")
# exp_group_object@anno$CancerType <- factor(exp_group_object@anno$CancerType, levels = time.point)
exp_group_object@anno$SegmentLabel <- factor(exp_group_object@anno$Prognosis)
exp_group_object <- epars_plot_setting(exp_group_object)
exp_group_object <- epars_heatmap_anno_data(exp_group_object)
ordersample <- rownames(dplyr::arrange(exp_group_object@anno, Prognosis))
p <- Heatmap(matrix = res$beta %>% scale(), 
        top_annotation = exp_group_object@plot_setting$heatmap_anno, 
        cluster_columns = T, #column_order = ordersample,
        heatmap_legend_param = list(title = "",legend_height = unit(4, "cm")),
        column_split = exp_group_object@plot_setting$group_df[, "Prognosis"]
        )
save_pdf(p, 
         filename = paste0(CachePath, 
                           "3.1 immune cells/different_prognositc/SpatialDecon.immcell.abundance.heatmap.pdf"), 
         height = 12, width = 12)

Cancertype.abundance <- res$beta %>% 
  as.data.frame() %>% 
  t() %>% 
  merge(exp_group_object@anno %>% dplyr::select(Prognosis), ., by = 0) %>% 
  column_to_rownames("Row.names") %>% 
  group_by(Prognosis) %>% 
  dplyr::summarise(dplyr::across(everything(), sum)) %>% 
  column_to_rownames("Prognosis") %>% 
  apply(1, \(.x) .x/sum(.x)) %>% 
  as.data.frame() %>% 
  rownames_to_column("CellType") %>% 
  pivot_longer(cols = -CellType, names_to = "Prognosis", values_to = "Cell.Proportion")

p <- Cancertype.abundance %>% 
  ggplot(aes(x = Prognosis, y = Cell.Proportion)) + 
  geom_col(aes(fill = CellType)) + 
  scale_fill_manual(values = col_vector) + 
  ylab("Cell Proportion")
ggsave(plot = p, 
       filename = paste0(CachePath,"3.1 immune cells/different_prognositc/SpatialDecon.immcell.proportion.pdf"), 
       height = 9, 
       width = 6)

plotdata <- res$beta %>% 
  as.data.frame() %>% 
  t() %>% 
  merge(exp_group_object@anno %>% dplyr::select(Prognosis), ., by = 0) #%>% 
  #dplyr::select(-monocytes.C)

plotdata$Prognosis <- plotdata$Prognosis %>% as.character()
source("~/script/Rscript/all.function.R")

adjusted_boxplot_nojitter(input_df = plotdata, 
                          id_colnames = colnames(plotdata)[1:2], 
                          x_aes = colnames(plotdata)[2], 
                          plot_dir = paste0(CachePath, "3.1 immune cells/different_prognositc/boxplot/"),
                          plot_suffix = ".abundance.boxplot.pdf", 
                          comparisons_list = combn(unique(plotdata$Prognosis), 2, simplify = F), 
                          x_order = c("bad", "good"), 
                          plot_width = 25, 
                          ncol = 6,
                          stat_method = "wilcox.test",
                          adjust_method = "BH")
```

### 3.3 Cibersort

为了探究M1-2的比例, 这里使用cibersort进行计算.

```{r cibersort}
source("~/script/Rscript/Cibersort/CIBERSORT.R")
Cibersort.result <- CIBERSORT(sig_matrix = readRDS("~/script/Rscript/Cibersort/LM22.rds"), 
                              mixture_file = exp_group_object@exp,
                              perm = 1000,
                              QN = FALSE, 
                              absolute = T)
write.xlsx(Cibersort.result %>% as.data.frame(), 
           file = paste0(CachePath, "3.3 Cibersort/Cibersort.res.absolut.xlsx"), rowNames = T)

Macrophages.cell <- Cibersort.result %>% 
  as.data.frame() %>% 
  dplyr::select(starts_with("Macrophages")) %>% 
  merge(., exp_group_object@anno %>% dplyr::select(CancerType), by = 0) %>% 
  dplyr::rename(SampleID = Row.names) %>% 
  dplyr::group_by(CancerType) %>% 
  dplyr::summarise(M0 = sum(`Macrophages M0`), 
            M1 = sum(`Macrophages M1`), 
            M2 = sum(`Macrophages M2`)) %>% 
  column_to_rownames("CancerType") %>% 
  apply(1, \(.x) .x/sum(.x)) %>% 
  t() %>% 
  as.data.frame() %>% 
  rownames_to_column("CancerType") %>% 
  pivot_longer(cols = -CancerType, names_to = "Macrophages", values_to = "Proprotion")

Macrophages.cell$CancerType <- factor(Macrophages.cell$CancerType, levels = time.point)

p <- Macrophages.cell %>% 
  ggplot(aes(x = CancerType , y = Proprotion)) + 
  geom_col(aes(fill = Macrophages))

ggsave(plot = p, filename = paste0(CachePath, "3.3 Cibersort/cibersort.Macrophages.Proportion .pdf"))
```

#### 3.3.1 M1-M2 proportion of different prognositc

```{r cibersort_macrophages}
Cibersort.result <- read.xlsx(xlsxFile = paste0(CachePath, "3.3 Cibersort/Cibersort.res.absolut.xlsx") ,
                              rowNames = T)
Macrophages.cell <- Cibersort.result %>% 
  as.data.frame() %>% 
  dplyr::select(starts_with("Macrophages")) %>% 
  merge(., exp_group_object@anno %>% dplyr::select(Prognosis), by = 0) %>% 
  dplyr::rename(SampleID = Row.names) %>% 
  dplyr::group_by(Prognosis) %>% 
  dplyr::summarise(M0 = sum(`Macrophages.M0`), 
            M1 = sum(`Macrophages.M1`), 
            M2 = sum(`Macrophages.M2`)) %>% 
  column_to_rownames("Prognosis") %>% 
  apply(1, \(.x) .x/sum(.x)) %>% 
  t() %>% 
  as.data.frame() %>% 
  rownames_to_column("Prognosis") %>% 
  pivot_longer(cols = -Prognosis, names_to = "Macrophages", values_to = "Proprotion")

p <- Macrophages.cell %>% 
  ggplot(aes(x = Prognosis , y = Proprotion)) + 
  geom_col(aes(fill = Macrophages))

ggsave(plot = p, 
       filename = paste0(CachePath, "3.3 Cibersort/Prognosis_cibersort.Macrophages.Proportion .pdf"),
       width = 6, height = 9)
```

## Part IV Macrophage ligand receptor analysis

利用Cellchat进行细胞配受体分析

```{r CellChat}
CachePath <- paste0(Work_canche_dir, "Part IV/")
PlotsPath <- paste0(Work_canche_dir, "Part IV/")
create_dir(path.list = c(CachePath, PlotsPath))
suppressPackageStartupMessages(library(CellChat))
suppressPackageStartupMessages(library(patchwork))
meta.data <- exp_group_object@anno %>% 
  dplyr::select(cell_type = Morphology) %>% 
  dplyr::filter(cell_type %in% c("M1", "M2", "TC"))

data.input <- exp_group_object@exp %>%  # data.input need to a martix
  as.data.frame() %>% 
  dplyr::select(rownames(meta.data)) %>% 
  as.matrix()

cellchat_martix <- createCellChat(object = data.input, meta = meta.data, group.by = "cell_type")
CellChatDB <- CellChatDB.human
cellchat_martix@DB <- CellChatDB
cellchat_martix <- subsetData(cellchat_martix)
cellchat_martix <- identifyOverExpressedGenes(cellchat_martix)
cellchat_martix <- identifyOverExpressedInteractions(cellchat_martix)
# execution.time = Sys.time() - ptm
# print(as.numeric(execution.time, units = "secs"))
# Compute the communication probability and infer cellular communication network
cellchat_martix <- computeCommunProb(cellchat_martix, type = "triMean")
cellchat_martix <- filterCommunication(cellchat_martix, min.cells = 10)
cellchat_martix <- computeCommunProbPathway(cellchat_martix)
cellchat_martix <- aggregateNet(cellchat_martix)
View(subsetCommunication(cellchat_martix))
saveRDS(cellchat_martix, file = paste0(CachePath, "/cellchat_martix.RDs"))

netAnalysis_contribution(cellchat_martix, signaling = "WNT")
netVisual_circle(cellchat_martix@net$count,  vertex.weight = as.numeric(table(cellchat_martix@idents)), weight.scale = T)
p <- netVisual_bubble(cellchat_martix, remove.isolate = F)
ggsave(paste0(CachePath, "/cellchat.res.dotplot.pdf"), height = 45, width = 9)
out.xlsx <- ggplot_build(p)
write.xlsx(out.xlsx$plot$data, file = paste0(CachePath, "/cellchat.res.xlsx"))
```

## Part V

### 5.1 Analysis of prognostic related markers

```{r Prognosis_DEG}
CachePath <- paste0(Work_canche_dir, "Part V/")
PlotsPath <- paste0(Work_canche_dir, "Part V/")
create_dir(path.list = c(CachePath, PlotsPath))

deg_group <- c("Prognosis(bad:good),Morphology$M2; Prognosis(bad:good),Morphology$M1; Prognosis(bad:good),Morphology$TC")
source("~/script/Rscript/Parallel script/library_EPARS_parallel.r")
exp_group_object@deg_groups <- deg_group
exp_group_object <- parallel_DEG(exp_group_object, CachePath = "./personal_analysis/rmd_Results/Part V/DEG/", core.num = 3)
parallel_volcano(exp_group_object, PlotsPath = "./personal_analysis/rmd_Results/Part V/DEG/volcano/")
parallel_heatmap(exp_group_object, CachePath = "personal_analysis/rmd_Results/Part V/DEG/heatmap/", PlotsPath = "personal_analysis/rmd_Results/Part V/DEG/heatmap/", core.num = 6)

### 预后基因的富集分析
exp_list <- exp_group_object@deg %>% 
  lapply(\(.x){.x@result_df}) %>% 
  lapply(\(.x){
    .x %>% dplyr::filter(change != "NOT")  %>% 
      split(f = .$change)
  }) %>% 
  .[sapply(., \(.x) length(.x) != 0)]

DEG_list <- list(M1_list =  exp_list[names(exp_list) %>% grep(patter = "M1", value = T)],
                 M2_list = exp_list[names(exp_list) %>% grep(patter = "M2", value = T)],
                 TC_list = exp_list[names(exp_list) %>% grep(patter = "TC", value = T)]) %>% 
  lapply(\(.x) {
    names(.x) <- names(.x) %>% str_extract(".*(?=\\,)")
    return(.x)
  })
enrichment.res_list <- enrichtment_tidydata(DEG_list, CachePath = paste0(CachePath, "DEG/Prognosis_enrichment_res/"))
plot.data <- enrichment.res_list %>% 
  purrr::map_depth(2, function(.x){
    do.call(rbind, .x) %>% 
      rownames_to_column("Group") %>% 
      mutate(Group = str_extract(Group, ".*(?=\\.)")) %>% 
      group_by(Group) %>% # 保存表格注释掉
      slice_min(p.adjust, n = 5) %>% # 这里取了每个组中矫正p值前5位 # 保存表格注释掉
      # 在画图之前, 先保存字符形式的Ratio
      mutate(GeneRatio = trans2num(GeneRatio), BgRatio = trans2num(BgRatio)) # 保存表格注释掉
  })
lapply(names(plot.data), \(.x){
  dir <- paste0(CachePath, "DEG/Prognosis_enrichment_res/",.x, "/")
  dir.create(dir, recursive = T, showWarnings = F)
  write.xlsx(plot.data[[.x]], file = paste0(dir, "enrichment.results.xlsx"))
})

lapply(names(plot.data), function(.x){
  lapply(names(plot.data[[.x]]), \(..x){
    print(..x)
    dir <- paste0(CachePath, "DEG/Prognosis_enrichment_res/",.x, "/", ..x, "/")
    dir.create(dir, recursive = T)
    p <- dotplot.function(plot.data[[.x]][[..x]])
    ggsave(filename = paste0(dir, "dotplot.pdf"), height = 20, width = 10, plot = p)
  })
})
```

```{r DEG_analysis_etc}
### 预后的差异基因
exp_group_object <- readRDS("~/work/DSP/YKKY0081天津胸科医院/DSP_WTA/personal_analysis/rmd_Results/Part V/DEG/exp_group_object.RDS")
exp_list <- exp_group_object@deg %>% 
  lapply(\(.x){.x@result_df}) %>% 
  lapply(\(.x){
    .x %>% dplyr::filter(change != "NOT") %>% rownames()
  }) %>% 
  .[sapply(., \(.x) length(.x) != 0)]
prognosis_DEG_list <- list(M1_list =  exp_list[names(exp_list) %>% grep(patter = "M1", value = T)],
                        M2_list = exp_list[names(exp_list) %>% grep(patter = "M2", value = T)],
                        TC_list = exp_list[names(exp_list) %>% grep(patter = "TC", value = T)]) %>% 
  lapply(\(.x) {
    names(.x) <- names(.x) %>% str_extract(".*(?=\\,)")
    return(.x)
  })
rm(exp_list)

### 前一步的差异基因

exp_group_object <- readRDS("./personal_analysis/rmd_Results/Part II/2.1 differential gene/DEG_res/exp_group_object.RDS")
exp_list <- exp_group_object@deg %>% 
  lapply(\(.x){.x@result_df}) %>% 
  lapply(\(.x){
    .x %>% dplyr::filter(change != "NOT") %>% rownames()
  }) %>% 
  .[sapply(., \(.x) length(.x) != 0)] %>% 
  .[!grepl("^Pro", names(.))]
cancer_DEG_list <- list(M1_list =  exp_list[names(exp_list) %>% grep(patter = "M1", value = T)],
                 M2_list = exp_list[names(exp_list) %>% grep(patter = "M2", value = T)],
                 TC_list = exp_list[names(exp_list) %>% grep(patter = "TC", value = T)]) %>% 
  lapply(\(.x) {
    names(.x) <- names(.x) %>% str_extract(".*(?=\\,)")
    return(.x)
  })
rm(exp_list)

# intersect(cancer_DEG_list$TC_list[[2]], prognosis_DEG_list$TC_list[[1]])

lapply(names(cancer_DEG_list$TC_list), \(.x) {
  intersect(cancer_DEG_list$TC_list[[.x]], prognosis_DEG_list$TC_list[[1]])
}) -> Tcell_intersect_gene
names(Tcell_intersect_gene) <- names(cancer_DEG_list$TC_list)


all_intersect_gene <- lapply(names(cancer_DEG_list), \(.x){
  ca <- cancer_DEG_list[[.x]]
  pro <- prognosis_DEG_list[[.x]]
  res <- lapply(names(ca), function(..x){
    intersect(ca[[..x]], pro[[1]])
  })
  names(res) <- names(ca)
  return(res)
})
names(all_intersect_gene) <- names(cancer_DEG_list)
all_intersect_gene
```

取交集后, 同预后差异基因有交集的主要集中在TC中, M1,M2细胞交集差异基因很少

```{r enrichment_insterect_gene}

all_instersect_enrichment_res <- 
  lapply(all_intersect_gene, \(.x){
  lapply(.x, \(..x){
    enrichKEGG.function(..x, qvalueCutoff = 1, pvalueCutoff = 1) %>% 
    as.data.frame()
  })
})
plot.data <- all_instersect_enrichment_res %>% 
  lapply(\(.x){
    .x <- .x %>% 
      do.call(rbind, .) %>% 
      rownames_to_column("Group") %>% 
      mutate(Group = str_extract(Group, ".*(?=\\.)")) %>% 
      group_by(Group) %>% # 保存表格注释掉
      slice_min(p.adjust, n = 5) %>% # 这里取了每个组中矫正p值前5位 # 保存表格注释掉
      # 在画图之前, 先保存字符形式的Ratio
      mutate(GeneRatio = trans2num(GeneRatio), BgRatio = trans2num(BgRatio)) # 保存表格注释掉
    return(.x)
  })
write.xlsx(plot.data, 
           file = paste0(CachePath, "enrichment_res/enrichment.results.xlsx"),
           overwrite = T)
pblapply(names(plot.data), function(.x){
  print(.x)
  dir <- paste0(CachePath, "enrichment_res/",.x, "/")
  dir.create(dir, recursive = T)
  p <- dotplot.function(plot.data[[.x]])
  ggsave(filename = paste0(dir, "dotplot.pdf"), height = 20, width = 10, plot = p)
})

```

### 5.2 The prognostic genes were verified

logrank test预后真正有影响的基因, 由于没有具体的生存状态, 这里使用DFS和是否复发进行生存分析

```{r WTA2Sample.expr}
anno <- dplyr::select(exp_group_object@anno, c(38:48),)
exp <- t(exp_group_object@exp)
meandata <- cbind(anno,exp)#将病人名称与AOI结合
meandata <- meandata %>% split(f = .[, "Morphology"])
#exp <- meandata %>% distinct(Sample,.keep_all = T)#随机选择相同病人的AOI
# exp <- aggregate(meandata[, 12:ncol(meandata)], by=list(meandata$TMA.position),mean) #计算同一病人不同AOI的均值
# rownames(exp) <- exp[,1]
# exp <- exp[,-1]
# exp_group_object_advanced@exp <- t(exp)
# exp_group_object_advanced@anno <- exp_group_object@anno[!duplicated(exp_group_object@anno$TMA.position),]
# rownames(exp_group_object_advanced@anno) <- exp_group_object_advanced@anno$TMA.position
# data <- exp_group_object_advanced
# data@anno$OS <- strsplit(data@anno$PFS_OS, " OS") %>% Reduce(rbind,.) %>% 
#   as.data.frame() %>% .[,2]
# data@anno$DFS <- strsplit(data@anno$PFS_OS, " OS") %>% Reduce(rbind,.) %>% 
#   as.data.frame() %>% .[,1]
# data@anno["08-1012-2",c("OS","DFS")] <- c("175","172.8")
# data@anno$OS <- data@anno$OS %>% gsub(pattern = " ",replacement = "")
# data@anno$DFS <- data@anno$DFS %>% gsub(pattern = "DFS ",replacement = "") %>% 
#   gsub(pattern = "DFS",replacement = "")
# data@anno["18-2116-4","DFS"] <- "54.8"
# data@anno$Status <- data@anno$Status %>% gsub(pattern = "未复发",replacement = "1") %>% 
#   gsub(pattern = "复发",replacement = "0") %>% as.numeric()
# data@anno$OS <- as.numeric(data@anno$OS)
# data@anno$DFS <- as.numeric(data@anno$DFS)
data <- lapply(meandata, \(.x){mean_Patient(.x)})
data[["ST"]]@anno
data[["ST"]]@anno <- data[["ST"]]@anno[1:4, ]
```

```{r Surv}
sig_gene <- pblapply(data, log_P, intersect_gene = all_intersect_gene) %>% 
  lapply(\(.x) append(.x, "SPP1")) %>% 
  .[sapply(., \(.x) length(.x) != 0)]
sig_gene$ST <- NULL
# exp <- data@exp %>% 
#   t() %>% 
#   as.data.frame() %>% 
#   dplyr::select(unlist(all_intersect_gene) %>% as.character() %>% unique())
# anno <- data@anno %>% 
#   dplyr::select(DFS, Status)
# cal_data <- merge(anno, exp, by = 0) %>% 
#   column_to_rownames("Row.names")
# phe <- cal_data[, 1:2]
# mySurv <- with(cal_data, Surv(DFS, Status))
# exprSet <- cal_data[, -c(1:3)] %>% t
# log_rank_p <- apply(exprSet, 1 , function(gene){
#   # gene=exprSet[1,]
#   phe$group <- ifelse(gene > median(gene), 'high', 'low')  
#   data.survdiff <- survdiff(mySurv ~ group, data = phe)
#   p.val = 1 - pchisq(data.survdiff$chisq, length(data.survdiff$n) - 1)
#   return(p.val)
# })
# log_rank_p <- sort(log_rank_p)
# table(log_rank_p < 0.05)
# sig_gene <- log_rank_p[log_rank_p < 0.05] %>% names()
```

利用logrank, 基因大于中位值的记为high, 低于记为low, 然后利用DFS和Status 构建的生存模型计算的, 挑选所有pvalue小于0.05的基因绘制曲线图

```{r survplot}
lapply(names(data), \(.x){
  dir <- paste0(CachePath, "survplot/", .x, "/")
  dir.create(dir, showWarnings = F)
  SurPlot(data, sig_gene, listnames = .x, CachePath = dir)
})



# lapply(names(sig_gene), \(.x){
#   gene <- sig_gene[[.x]]
#   dir <- paste0(CachePath, "survplot/", .x, "/")
#   dir.create(dir, recursive = T, showWarnings = F)
#   for (g in gene) {
#   phe <- phe %>%
#     dplyr::mutate(!!sym(g) := ifelse(exprSet[g, ] > median(exprSet[g, ]),'high','low'))
#   sfit1 <- survfit(as.formula(paste0("Surv(DFS, Status) ~ ", g)), data = phe)
#   p <- ggsurvplot(sfit1, pval =TRUE, data = phe, risk.table = TRUE)
#   pdf(file = paste0(dir, g, "_survplot.pdf"), onefile = F)
#   print(p)
#   dev.off()
# }
# })
# 
# for (g in sig_gene) {
#   phe <- phe %>%
#     dplyr::mutate(!!sym(g) := ifelse(exprSet[g, ] > median(exprSet[g, ]),'high','low'))
#   sfit1 <- survfit(as.formula(paste0("Surv(DFS, Status) ~ ", g)), data = phe)
#   p <- ggsurvplot(sfit1, pval =TRUE, data = phe, risk.table = TRUE)
#   pdf(file = paste0(CachePath, "survplot/", g, "_survplot.pdf"), onefile = F)
#   print(p)
#   dev.off()
# }
```

### 5.3 Cell interactions associated with prognosis

```{r CellChat2}
meta.data_all <- exp_group_object@anno %>%
dplyr::select(cell_type = Morphology, Prognosis) %>%
dplyr::filter(cell_type %in% c("M1", "M2", "TC"))
data.input_all <- exp_group_object@exp %>%  # data.input need to a martix
as.data.frame() %>%
dplyr::select(rownames(meta.data_all))
#### input good ###
data.input <- data.input_all %>%
as.data.frame() %>%
dplyr::select(rownames(dplyr::filter(meta.data_all, Prognosis == "good"))) %>%
as.matrix()
meta.data <- meta.data_all %>%
dplyr::filter(Prognosis == "good") %>%
dplyr::select(-Prognosis)
cellchat_martix <- createCellChat(object = data.input, meta = meta.data, group.by = "cell_type")
CellChatDB <- CellChatDB.human
cellchat_martix@DB <- CellChatDB
cellchat_martix <- subsetData(cellchat_martix)
cellchat_martix <- identifyOverExpressedGenes(cellchat_martix)
cellchat_martix <- identifyOverExpressedInteractions(cellchat_martix)
# execution.time = Sys.time() - ptm
# print(as.numeric(execution.time, units = "secs"))
# Compute the communication probability and infer cellular communication network
cellchat_martix <- computeCommunProb(cellchat_martix, type = "triMean")
cellchat_martix <- filterCommunication(cellchat_martix, min.cells = 10)
cellchat_martix <- computeCommunProbPathway(cellchat_martix)
cellchat_martix <- aggregateNet(cellchat_martix)
# View(subsetCommunication(cellchat_martix))
saveRDS(cellchat_martix, file = paste0(CachePath, "5.3 cellchat/5.3_good_cellchat_martix.RDs"))
netAnalysis_contribution(cellchat_martix, signaling = "WNT")
netVisual_circle(cellchat_martix@net$count,  vertex.weight = as.numeric(table(cellchat_martix@idents)), weight.scale = T)
p <- netVisual_bubble(cellchat_martix, remove.isolate = F)
ggsave(paste0(CachePath, "5.3 cellchat/5.3_good_cellchat.res.dotplot.pdf"), height = 45, width = 9)
out.xlsx <- ggplot_build(p)
write.xlsx(out.xlsx$plot$data, file = paste0(CachePath, "5.3 cellchat/5.3_good_cellchat.res.xlsx"))
rm(data.input)
CachePath <- paste0(Work_canche_dir, "Part V/")
PlotsPath <- paste0(Work_canche_dir, "Part V/")
create_dir(path.list = c(CachePath, PlotsPath))
# default parameter
Work_canche_dir <- "./personal_analysis/rmd_Results/"
Work_plot_dir<- "./personal_analysis/rmd_Results/"
CachePath <- paste0(Work_canche_dir, "Part V/")
PlotsPath <- paste0(Work_canche_dir, "Part V/")
saveRDS(cellchat_martix, file = paste0(CachePath, "5.3 cellchat/5.3_good_cellchat_martix.RDs"))
netAnalysis_contribution(cellchat_martix, signaling = "WNT")
netVisual_circle(cellchat_martix@net$count,  vertex.weight = as.numeric(table(cellchat_martix@idents)), weight.scale = T)
p <- netVisual_bubble(cellchat_martix, remove.isolate = F)
ggsave(paste0(CachePath, "5.3 cellchat/5.3_good_cellchat.res.dotplot.pdf"), height = 45, width = 9)
out.xlsx <- ggplot_build(p)
write.xlsx(out.xlsx$plot$data, file = paste0(CachePath, "5.3 cellchat/5.3_good_cellchat.res.xlsx"))
rm(data.input)
#### bad ###
data.input <- data.input_all %>%
as.data.frame() %>%
dplyr::select(rownames(dplyr::filter(meta.data_all, Prognosis == "bad"))) %>%
as.matrix()
meta.data <- meta.data_all %>%
dplyr::filter(Prognosis == "bad") %>%
dplyr::select(-Prognosis)
cellchat_martix <- createCellChat(object = data.input, meta = meta.data, group.by = "cell_type")
CellChatDB <- CellChatDB.human
cellchat_martix@DB <- CellChatDB
cellchat_martix <- subsetData(cellchat_martix)
cellchat_martix <- identifyOverExpressedGenes(cellchat_martix)
cellchat_martix <- identifyOverExpressedInteractions(cellchat_martix)
# execution.time = Sys.time() - ptm
# print(as.numeric(execution.time, units = "secs"))
# Compute the communication probability and infer cellular communication network
cellchat_martix <- computeCommunProb(cellchat_martix, type = "triMean")
cellchat_martix <- filterCommunication(cellchat_martix, min.cells = 10)
cellchat_martix <- computeCommunProbPathway(cellchat_martix)
cellchat_martix <- aggregateNet(cellchat_martix)
# View(subsetCommunication(cellchat_martix))
saveRDS(cellchat_martix, file = paste0(CachePath, "5.3 cellchat/5.3_bad_cellchat_martix.RDs"))
netAnalysis_contribution(cellchat_martix, signaling = "WNT")
netVisual_circle(cellchat_martix@net$count,  vertex.weight = as.numeric(table(cellchat_martix@idents)), weight.scale = T)
p <- netVisual_bubble(cellchat_martix, remove.isolate = F)
ggsave(paste0(CachePath, "5.3 cellchat/5.3_bad_cellchat.res.dotplot.pdf"), height = 45, width = 9)
out.xlsx <- ggplot_build(p)
write.xlsx(out.xlsx$plot$data, file = paste0(CachePath, "5.3 cellchat/5.3_bad_cellchat.res.xlsx"))

bad_res <- out.xlsx$plot$data
good_res <- read.xlsx(paste0(CachePath, "5.3 cellchat/5.3_good_cellchat.res.xlsx"))
common_type <- intersect(good_res$source.target, bad_res$source.target)

bad_res <- dplyr::filter(bad_res, source.target  %in% common_type) %>% 
  dplyr::mutate(Group = "bad") %>% 
  dplyr::select(Group, starts_with("inter"), source.target, prob, pval)
good_res <- dplyr::filter(good_res, source.target  %in% common_type) %>% 
  dplyr::mutate(Group = "good") %>% 
  dplyr::select(Group, starts_with("inter"), source.target, prob, pval)
plotdata <- rbind(bad_res, good_res) %>% 
  mutate(Group = paste0(source.target, "_", Group))


color.use <- RColorBrewer::brewer.pal(n = 10, name = "Spectral")
p <- plotdata %>% 
  ggplot(aes(x = Group, y = interaction_name_2, color = prob)) +
  geom_point(pch = 16) +
  theme_linedraw() + 
  theme(panel.grid.major = element_line(color = "grey90"),
        panel.grid.minor = element_line(color = "grey90")) +
  scale_color_gradientn(colors =  colorRampPalette(color.use)(99), 
                         na.value = "white", 
                         limits = c(quantile(plotdata$prob, 0, na.rm = T), 
                                    quantile(plotdata$prob, 1, na.rm = T)),  
                         breaks = c(quantile(plotdata$prob, 0, na.rm = T), 
                                    quantile(plotdata$prob, 1, na.rm = T)), 
                         labels = c("min", "max")) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  xlab("") + 
  ylab("")
ggsave(plot = p, filename = paste0(CachePath, "5.3 cellchat/commom_type_res.pdf"), height = 45, width = 9)
  
```

#### 5.3.2 Multiple groups are analyzed

这里利用官方对于不同数据集分析的教程进行尝试

```{r allmetadata}
meta.data_all <- exp_group_object@anno %>%
dplyr::select(cell_type = Morphology, Prognosis) %>%
dplyr::filter(cell_type %in% c("M1", "M2", "TC"))
data.input_all <- exp_group_object@exp %>%  # data.input need to a martix
as.data.frame() %>%
dplyr::select(rownames(meta.data_all))
```

```{r good_group}
#### input good ####
data.input <- data.input_all %>%
as.data.frame() %>%
dplyr::select(rownames(dplyr::filter(meta.data_all, Prognosis == "good"))) %>%
as.matrix()
meta.data <- meta.data_all %>%
dplyr::filter(Prognosis == "good") %>%
dplyr::select(-Prognosis)
cellchat_martix_good <- createCellChat(object = data.input, meta = meta.data, group.by = "cell_type")
CellChatDB <- CellChatDB.human
cellchat_martix_good@DB <- CellChatDB
cellchat_martix_good <- subsetData(cellchat_martix_good)
cellchat_martix_good <- identifyOverExpressedGenes(cellchat_martix_good)
cellchat_martix_good <- identifyOverExpressedInteractions(cellchat_martix_good)
# execution.time = Sys.time() - ptm
# print(as.numeric(execution.time, units = "secs"))
# Compute the communication probability and infer cellular communication network
cellchat_martix_good <- computeCommunProb(cellchat_martix_good, type = "triMean")
cellchat_martix_good <- filterCommunication(cellchat_martix_good, min.cells = 10)
cellchat_martix_good <- computeCommunProbPathway(cellchat_martix_good)
cellchat_martix_good <- aggregateNet(cellchat_martix_good)
# View(subsetCommunication(cellchat_martix_good))
ouptfile <- subsetCommunication(cellchat_martix_good)
write.xlsx(ouptfile, file = paste0(CachePath, "5.3 cellchat/Multiple_groups/5.3_good_cellchat.res.xlsx"))
saveRDS(cellchat_martix_good, file = paste0(CachePath, "5.3 cellchat/Multiple_groups/5.3_good_cellchat_martix.RDs"))
# netAnalysis_contribution(cellchat_martix_good, signaling = "WNT")
pdf(file = paste0(CachePath, "5.3 cellchat/Multiple_groups/5.3_good_cellchat.res.circle.pdf"))
p <- netVisual_circle(cellchat_martix_good@net$count,  vertex.weight = as.numeric(table(cellchat_martix_good@idents)), weight.scale = T)
dev.off()
p <- netVisual_bubble(cellchat_martix_good, remove.isolate = F)
ggsave(plot = p, paste0(CachePath, "5.3 cellchat/Multiple_groups/5.3_good_cellchat.res.dotplot.pdf"), height = 45, width = 9)
# out.xlsx <- ggplot_build(p)
# write.xlsx(out.xlsx$plot$data, file = paste0(CachePath, "5.3 cellchat/Multiple_groups/5.3_good_cellchat.res.xlsx"))
```

```{r bad_group}
data.input <- data.input_all %>%
as.data.frame() %>%
dplyr::select(rownames(dplyr::filter(meta.data_all, Prognosis == "bad"))) %>%
as.matrix()
meta.data <- meta.data_all %>%
dplyr::filter(Prognosis == "bad") %>%
dplyr::select(-Prognosis)
cellchat_martix_bad <- createCellChat(object = data.input, meta = meta.data, group.by = "cell_type")

CellChatDB <- CellChatDB.human
cellchat_martix_bad@DB <- CellChatDB
cellchat_martix_bad <- subsetData(cellchat_martix_bad)
cellchat_martix_bad <- identifyOverExpressedGenes(cellchat_martix_bad)
cellchat_martix_bad <- identifyOverExpressedInteractions(cellchat_martix_bad)
# execution.time = Sys.time() - ptm
# print(as.numeric(execution.time, units = "secs"))
# Compute the communication probability and infer cellular communication network
cellchat_martix_bad <- computeCommunProb(cellchat_martix_bad, type = "triMean")
cellchat_martix_bad <- filterCommunication(cellchat_martix_bad, min.cells = 10)
cellchat_martix_bad <- computeCommunProbPathway(cellchat_martix_bad)
cellchat_martix_bad <- aggregateNet(cellchat_martix_bad)
# View(subsetCommunication(cellchat_martix_bad))
ouptfile <- subsetCommunication(cellchat_martix_bad)
write.xlsx(ouptfile, file = paste0(CachePath, "5.3 cellchat/Multiple_groups/5.3_bad_cellchat.res.xlsx"))
saveRDS(cellchat_martix_bad, file = paste0(CachePath, "5.3 cellchat/Multiple_groups/5.3_bad_cellchat_martix.RDs"))
# netAnalysis_contribution(cellchat_martix_bad, signaling = "WNT")
pdf(file = paste0(CachePath, "5.3 cellchat/Multiple_groups/5.3_bad_cellchat.res.circle.pdf"))
p <- netVisual_circle(cellchat_martix_bad@net$count,  vertex.weight = as.numeric(table(cellchat_martix_bad@idents)), weight.scale = T)
p
dev.off()
p <- netVisual_bubble(cellchat_martix_bad, remove.isolate = F)
ggsave(plot = p, paste0(CachePath, "5.3 cellchat/Multiple_groups/5.3_bad_cellchat.res.dotplot.pdf"), height = 45, width = 9)
# out.xlsx <- ggplot_build(p)
# write.xlsx(out.xlsx$plot$data, file = paste0(CachePath, "5.3 cellchat/Multiple_groups/5.3_bad_cellchat.res.xlsx"))

```

```{r chellchat_Multiple}
#### merge the cellchat object
object.list <- list(Good = cellchat_martix_good, Bad = cellchat_martix_bad)
cellchat <- mergeCellChat(object.list, add.names = names(object.list))
saveRDS(object.list, file = paste0(CachePath, "5.3 cellchat/Multiple_groups/ellchat_object.list_humanSkin_NL_LS.RDs"))
saveRDS(cellchat, file = paste0(CachePath, "5.3 cellchat/Multiple_groups/cellchat_merged_humanSkin_NL_LS.RDs"))
```

###### **5.3.2.1 Compare the total number of interactions and interaction strength**

To answer the question on whether the cell-cell communication is enhanced or not, CellChat compares the total number of interactions and interaction strength of the inferred cell-cell communication networks from different biological conditions.

```{r 5.3.2.1}
gg1 <- compareInteractions(cellchat, show.legend = F, group = c(1,2))
gg2 <- compareInteractions(cellchat, show.legend = F, group = c(1,2), measure = "weight")
p <- gg1 + gg2
ggsave(plot = p, filename = paste0(CachePath, "5.3 cellchat/Multiple_groups/number and strength barplot.pdf"))
```

###### 5.3.2.2

Circle plot showing differential number of interactions or interaction strength among different cell populations across two datasets

The differential number of interactions or interaction strength in the cell-cell communication network between two datasets can be visualized using circle plot, where redred (or blueblue) colored edges represent increasedincreased (or decreaseddecreased) signaling in the second dataset compared to the first one.

```{r 5.3.2.2}
par(mfrow = c(1,2), xpd=TRUE)
pdf(file = paste0(CachePath, "5.3 cellchat/Multiple_groups/Differential number of interactions.pdf"))
gg1 <- netVisual_diffInteraction(cellchat, weight.scale = T)
gg1
dev.off()
pdf(file = paste0(CachePath, "5.3 cellchat/Multiple_groups/Differential strength of interactions.pdf"))
gg2 <- netVisual_diffInteraction(cellchat, weight.scale = T, measure = "weight")
gg2
dev.off()
```

###### 5.3.2.3

```{r 5.3.2.3}
gg1 <- netVisual_heatmap(cellchat)
#> Do heatmap based on a merged object
gg2 <- netVisual_heatmap(cellchat, measure = "weight")
#> Do heatmap based on a merged object
p <- gg1 + gg2
pdf(file = paste0(CachePath, 
                  "5.3 cellchat/Multiple_groups/Differential strength(number) of interactions.pdf"), width = 12)
p
dev.off()

```

## Part VI

### Find all marker in the CancerType

```{r find_marker}
exp_group_object_VI <- DSP_FindAllMarker(exp_group_object, 
                                         ident_col = "CancerType", 
                                         CachePath = "./personal_analysis/rmd_Results/Part VI/")
exp_group_object_VI@anno$Morphology <- factor(exp_group_object_VI@anno$Morphology, 
                                              levels = c("M1", "M2", "TC", "ST"))
parallel_heatmap(exp_group_object_VI, CachePath = "./personal_analysis/rmd_Results/Part VI/", PlotsPath = "./personal_analysis/rmd_Results/Part VI/Heatmap/", core.num = 1, top = 20, order_str = "Morphology")

DEGlist <- exp_group_object_VI@deg %>% 
  lapply(\(.x){.x@result_df}) %>% 
  lapply(\(.x){
    .x %>% dplyr::filter(change == "UP")  %>% 
      slice_min(FDR, n = 10) %>% 
      rownames()
  }) %>% 
  .[sapply(., \(.x) length(.x) != 0)] %>% 
  unlist() %>% 
  unique()
exp <- exp_group_object@exp %>% 
  as.data.frame() %>% 
  dplyr::filter(rownames(.)  %in% DEGlist)
anno <- exp_group_object@anno %>% 
  dplyr::select(CancerType, Morphology, Prognosis)

# plotdata <- merge(anno, t(exp), by = 0, sort = F)
anno$CancerType <- factor(anno$CancerType, levels = time.point)
order_sample <- rownames(arrange(anno, Morphology))
ha <- HeatmapAnnotation(df = anno)
p <- Heatmap(scale(exp), 
        top_annotation = ha, 
        show_column_names = F, 
        cluster_columns = F, 
        column_order = order_sample, 
        km = 2, 
        column_split = anno$CancerType, 
        border = T)
library(tidyHeatmap)
save_pdf(p, filename = "./personal_analysis/rmd_Results/Part VI/all_DEGs_Heatmap.pdf", height = 12, width = 12)
```
